Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-19 writeup/gz_hubble_data.tex
Better use and definition of 'bias'. Closes #71.

To facilitate correction of classifications in the presence of known biasing
effects, GZH includes two different samples of simulated \hst{} images: real
\hst{} galaxy images (Section \ref{ssec:legacy_surveys}) with nuclear emission
added to simulate AGN host galaxies, and lower-redshift galaxy images from SDSS
(Section \ref{ssec:sdss}), artificially redshifted from $0.3 \leq z \leq 1$.
Each is described separately below.

@@

Images with simulated AGN were classified in the interface in an identical
manner and evenly distributed with unaltered images of the galaxies.
Classifiers were not explicitly told that the images had been altered during
classification, as the goal was to measure the effect on normal classifications
in as unbiased a manner as possible. Following classification, a classifier
could view a page with additional details about each galaxy they had
classified; where applicable these pages contained further information
regarding image modifications. \citet{sim14} uses similarly-processed
images for reliable identification of bulgeless galaxies using SDSS data.

@@

The final number of \ferengi{} images produced for each galaxy is ultimately a
function of galaxy's redshift (since the new images cannot be resampled at
better angular resolution than the original SDSS data), as well as the number
of $e$ values selected. Table~\ref{tbl:ferengivalues} summarizes the redshifted
images produced for GZH. The use of these images to correct for redshift-dependent
classification bias is described in Section \ref{sec:debiasing}.

@@

Redshifts for the GZH catalog were compiled from a variety of sources. For each
galaxy, the primary redshift used for debiasing calculations
(Section~\ref{sec:debiasing}) is in the $\tt Z\_BEST$ column of
Table~\ref{tbl:catalog_hst}. The redshift type (spectroscopic: $\tt SPEC\_Z$,
photometric: $\tt PHOTO\_Z$, or grism: $ \tt GRISM\_Z$) is listed in the column
$\tt Z\_BEST\_TYPE$, and the source catalog of the redshift is included as $\tt
Z\_BEST\_SOURCE$.

@@

The previous versions of Galaxy Zoo morphology classifications
\citep{lin11,wil13} were based on observations of galaxies in the Sloan Digital
Sky Survey (SDSS) which have a median redshift of $z<0.2$. In these cases, it
was assumed that there was no cosmological evolution of the morphologies of
galaxies and therefore any observed changes in the distribution of galaxies
with different consensus morphologies was due to the effects of redshift on the
image quality (\ie, the reduction in physical resolution, surface brightness
dimming, etc). For both previous releases of Galaxy~Zoo morphologies, a correction for
redshift-dependent bias was applied based on matching the classification
fractions at the highest redshifts with those at the lowest redshift.
\citet{bam09} and \citet{wil13} provide complete descriptions of the process
for GZ1 and GZ2, respectively.

In the GZH samples, the redshift range is large enough that cosmological
evolution of the types and morphologies of galaxies is expected for galaxies in
the sample. As a result, the previous methods of correcting for redshift
dependent bias do not work. In addition, the effects of band shifting will
change the images even more across these redshift ranges. 
%Figure~\ref{fig:exampleFERENGI} illustrates some of the possible effects of
%losing features in spiral galaxies at high redshift.

@@

The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the image as seen on screen.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. There may also be other parameters that affect
classifier accuracy, such as the proximity of close companions or bias as a
function of individual classifier ability and skill. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters intended to capture the bulk of the change in bias: the
$r^\prime$-band surface brightness ($\mu_r$; intrinsic) and redshift ($z$;
extrinsic).

The change in bias as a function of $\mu_r$ and \zsim{} is measured using the
\ferengi{} images over all the evolutionary correction factors. It is assumed
that the ``true'' (ie, debiased) vote fraction $f_{\mu,z}$ for an
artificially-redshifted galaxy can be expressed as:

@@

The objective is to use these data to predict, for a galaxy with a measured
\ffeaturesz{} value, what its \ffeatures{} value \emph{would have been} if it
had been viewed at $z=0.3$. This predicted value is defined as the debiased
vote fraction \ffeaturesdebiased, and is calculated by applying a correction to
the measured value of \ffeatures, determined by the $\zeta$ function described
in the previous section. A reliable predicted value can be obtained so long as
the relationship between \ffeaturesz{} and \ffeaturesrest{} is single-valued;
that is, for a given \ffeaturesz, there is exactly one corresponding value of
\ffeatures{} at $z=0.3$.

@@

Few galaxies in the sample have sufficiently high corrections to completely
change them from being confidently smooth to featured following the bias
correction (Figure~\ref{fig:debiased_corrections}). As a check, highly boosted
galaxies are compared to the expert classifications in CANDELS \citep{kar15}.
There are only eight galaxies that were strongly boosted (\ffeatures$<0.2$ and
\fbest$>0.5$) in GZH and also appears in the CANDELS expert sample. Of those
eight galaxies, \citet{kar15} classify 5 as spheroids, 2 as disks, and 1 as
irregular/disky. The \fbest{} values for GZH are all between 0.5 and 0.6,
making them intermediate disk candidates. One of the spheroids has a hint of
extended emission to the south that may have been missed by the CANDELS
experts, but almost all images are dim and with very low surface brightnesses
in the ACS imaging, making positive identification challenging. Since the
surveys use different rest-frame filters and the overlap sample is tiny,
though, detailed comparisons between the overall morphologies are highly
difficult.

@@

The GZH catalog is corrected for classification bias only for the first
question of the GZH decision tree (Section~\ref{sec:debiasing}), which asks
{\it ``Is the galaxy smooth and round, with no sign of a disk?''} For this
question, the catalog provides the additional parameters $\tt debiased$, $\tt
lower~limit$, $\tt upper~limit$, and $\tt best$ vote fractions. The $\tt best$
fraction for \ffeatures{} is chosen based on the categorization of the galaxy:
if it is ``correctable'', $\tt best = debiased$; if it is a lower limit, $\tt
best = lower~limit$; if neither condition applies, then $\tt best = weighted$.

@@

Finally, there are 7,681~galaxies in the GOODS-S field with morphological
classifications in both GZH and the Galaxy~Zoo:~CANDELS project (Simmons
et~al., submitted). Since both the sensitivity and filters for the two sets of
images differ (and there is no debiased correction applied to GZC), there is no
prior reason to expect a perfect correlation between the separate vote
fractions for the projects. Briefly, we note that the \ffeatures{} value for
GZH is on average higher than GZC; the effect is strongest at
$f_\mathrm{features,GZC}<0.3$, for which roughly half the galaxies have
$f_\mathrm{features,GZH}>0.5$. However, the correlation between vote fractions
is single-valued (although not linear, with a Pearson correlation coefficient
of $r=0.6$), and should be possible to calibrate using a similar approach to
that described in Section~\ref{sec:debiasing}; the correlation between other
tasks, such as edge-on galaxies is significantly stronger ($r=0.9$). So while
the raw vote fractions are not directly comparable, the initial analysis
indicates that the broad morphologies are at least consistent.

>>>>

To facilitate correction of classifications in the presence of known redshift
bias (Section~\ref{sec:debiasing}), GZH includes two different samples of
simulated \hst{} images: real \hst{} galaxy images (Section
\ref{ssec:legacy_surveys}) with nuclear emission added to simulate AGN host
galaxies, and lower-redshift galaxy images from SDSS (Section \ref{ssec:sdss}),
artificially redshifted from $0.3 \leq z \leq 1$.  Each is described separately
below.

@@

Images with simulated AGN were classified in the interface in an identical
manner and evenly distributed with unaltered images of the galaxies.
Volunteers were not explicitly told that the images had been altered during
classification, as the goal was to measure the effect on normal classifications
usng the same technique as closely as possible. Following classification, a
classifier could view a page with additional details about each galaxy; where
applicable, these pages contained further information regarding image
modifications. \citet{sim14} uses similarly-processed images for reliable
identification of bulgeless galaxies using SDSS data.

@@

The final number of \ferengi{} images produced for each galaxy is ultimately a
function of galaxy's redshift (since the new images cannot be resampled at
better angular resolution than the original SDSS data), as well as the number
of $e$ values selected. Table~\ref{tbl:ferengivalues} summarizes the redshifted
images produced for GZH. The use of these images to correct for redshift-dependent
bias is described in Section \ref{sec:debiasing}.

@@

Redshifts for the GZH catalog were compiled from a variety of sources. For each
galaxy, the primary redshift is in the $\tt Z\_BEST$ column of
Table~\ref{tbl:catalog_hst}. The redshift type (spectroscopic: $\tt SPEC\_Z$,
photometric: $\tt PHOTO\_Z$, or grism: $ \tt GRISM\_Z$) is listed in the column
$\tt Z\_BEST\_TYPE$, and the source catalog of the redshift is included as $\tt
Z\_BEST\_SOURCE$.

@@

The previous versions of Galaxy Zoo morphology classifications
\citep{lin11,wil13} were based on observations of galaxies in the Sloan Digital
Sky Survey (SDSS) which have a median redshift of $z<0.2$. In these cases, it
was assumed that there was no cosmological evolution of the morphologies of
galaxies and therefore any observed changes in the morphological distribution
were due to a redshift-dependent bias that affects image quality and
classification accuracy (ie, galaxies of a given mass/size will appear smaller
and dimmer at higher redshifts).  This bias is not unique to the crowdsourcing
technique; its dependence in data quality is a potential problem for
\emph{both} automated and visual classifications, and must be addressed in
order to accurately measure demographics over any significant redshift range.

For both previous releases of Galaxy~Zoo morphologies, a correction for
redshift-dependent bias was applied based on matching the classification
fractions at the highest redshifts with those at the lowest redshift.
\citet{bam09} and \citet{wil13} provide complete descriptions of the process
for GZ1 and GZ2, respectively.

In the GZH samples, the redshift range is large enough that cosmological
evolution of the types and morphologies of galaxies is expected for galaxies in
the sample. As a result, the previous methods of correcting for
redshift-dependent bias do not work. In addition, the effects of band shifting
will change the images even more across these redshift ranges. 
%Figure~\ref{fig:exampleFERENGI} illustrates some of the possible effects of
%losing features in spiral galaxies at high redshift.

@@

The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the image as seen on screen.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. There may also be other parameters that affect
classifier accuracy, such as the proximity of close companions or 
individual classifier ability and skill. The combination of all
such parameters forms a high-dimensional space, and it is not clear how to
separate this into individual effects. Instead, the method used here employs
only two parameters intended to capture the bulk of the change in bias: the
$r^\prime$-band surface brightness ($\mu_r$; intrinsic) and redshift ($z$;
extrinsic).

This dependence of the vote fraction on $\mu_r$ and \zsim{} is measured using the
\ferengi{} images over all the evolutionary correction factors. It is assumed
that the ``true'' (ie, debiased) vote fraction $f_{\mu,z}$ for an
artificially-redshifted galaxy can be expressed as:

@@

The objective is to use these data to predict, for a galaxy with a measured
\ffeaturesz{} value, what its \ffeatures{} value \emph{would have been} if it
had been viewed at $z=0.3$. This predicted value is defined as the ``debiased''
vote fraction \ffeaturesdebiased, and is calculated by applying a correction to
the measured value of \ffeatures, determined by the $\zeta$ function described
in the previous section. A reliable predicted value can be obtained so long as
the relationship between \ffeaturesz{} and \ffeaturesrest{} is single-valued;
that is, for a given \ffeaturesz, there is exactly one corresponding value of
\ffeatures{} at $z=0.3$.

@@

Few galaxies in the sample have sufficiently high corrections to completely
change them from being confidently ``smooth'' to ``featured'' following the
correction for redshift bias (Figure~\ref{fig:debiased_corrections}). As a
check, highly boosted galaxies are compared to the expert classifications in
CANDELS \citep{kar15}.  There are only eight galaxies that were strongly
boosted (\ffeatures$<0.2$ and \fbest$>0.5$) in GZH and also appears in the
CANDELS expert sample. Of those eight galaxies, \citet{kar15} classify 5 as
spheroids, 2 as disks, and 1 as irregular/disky. The \fbest{} values for GZH
are all between 0.5 and 0.6, making them intermediate disk candidates. One of
the spheroids has a hint of extended emission to the south that may have been
missed by the CANDELS experts, but almost all images are dim and with very low
surface brightnesses in the ACS imaging, making positive identification
challenging. Since the surveys use different rest-frame filters and the overlap
sample is tiny, though, detailed comparisons between the overall morphologies
are highly difficult.

@@

The GZH catalog is corrected for redshift bias only for the first
question of the GZH decision tree (Section~\ref{sec:debiasing}), which asks
{\it ``Is the galaxy smooth and round, with no sign of a disk?''} For this
question, the catalog provides the additional parameters $\tt debiased$, $\tt
lower~limit$, $\tt upper~limit$, and $\tt best$ vote fractions. The $\tt best$
fraction for \ffeatures{} is chosen based on the categorization of the galaxy:
if it is ``correctable'', $\tt best = debiased$; if it is a lower limit, $\tt
best = lower~limit$; if neither condition applies, then $\tt best = weighted$.

@@

Finally, there are 7,681~galaxies in the GOODS-S field with morphological
classifications in both GZH and the Galaxy~Zoo:~CANDELS project (Simmons
et~al., submitted). Since both the sensitivity and filters for the two sets of
images differ (and there is no correction for redshift bias applied to GZC), there is no
prior reason to expect a perfect correlation between the separate vote
fractions for the projects. Briefly, we note that the \ffeatures{} value for
GZH is on average higher than GZC; the effect is strongest at
$f_\mathrm{features,GZC}<0.3$, for which roughly half the galaxies have
$f_\mathrm{features,GZH}>0.5$. However, the correlation between vote fractions
is single-valued (although not linear, with a Pearson correlation coefficient
of $r=0.6$), and should be possible to calibrate using a similar approach to
that described in Section~\ref{sec:debiasing}; the correlation between other
tasks, such as edge-on galaxies is significantly stronger ($r=0.9$). So while
the raw vote fractions are not directly comparable, the initial analysis
indicates that the broad morphologies are at least consistent.
